A Plasma Torus Around a Young Low-Mass Star

### 2025-03-13 paper/ms.tex
update

{\it Age: Isochrones}---The color and absolute magnitude of
TIC~141146667 suggest that it is a pre-main-sequence M dwarf, similar
to all other known CPVs \cite{Stauffer2017,Stauffer2021,Bouma2024}.
The star's proximity ($d$=58\,pc) and its high galactic latitude
($b$=$+$53$^\circ$) yield negligible interstellar reddening along the
line of sight \cite{Green2019}.
{\bf Figure~X} shows the location of the star in the color--absolute
magnitude diagram (CAMD)
relative to young stellar populations including Upper Scorpius (USco),
Upper-Centaurus-Lupus (UCL), IC~2602, and the Pleiades.
To make this diagram, {\bf TODO}we adopted the USco and UCL members
from X, the IC~2602 members from Y, and the Pleiades members from Z,
and applied the same reddening corrections as in \cite{Bouma2022}.
To measure the isochrone age of TIC~141146667, we then followed the 
empirical approach proposed by \cite{Gagne2020} and
implemented in \cite{Bouma2022}.
The idea is to estimate the star (or star cluster's) age by
interpolating between the observed isochrones of star clusters with
ages that have been otherwise calibrated, for instance using
the lithium depletion-boundary method.
While this circumvents known modelling issues in molecular opacities
and starspot coverage for young M dwarfs,
the ages returned by this procedure depend on the ages
assumed for each reference cluster.
In this case, the most important reference is UCL, since TIC~141146667
overlaps directly on its isochrone on the CAMD.
We adopted the 16~Myr age for UCL from \cite{PecautMamajek2016}, which
is consistent with ages determined by more recent sub-clustering
analyses \cite{Ratzenbock2023} for the largest sub-populations.
This in turns yields an isochronal age for TIC~141146667 of
$t_{\rm iso}$=$16^{+19}_{-6}$\,Myr.
Qualitatively, this range of uncertainty is consistent with no stars
in the $\approx$40\,Myr IC~2602 cluster with similar redness as
TIC~141146667 being equally bright.

{\it Age: Lithium}--- Frankly, it's not obviously there!  (but maybe
this is bc of the blending, and a pEW would show smth small)
%TODO TODO TODO

>>>>

{\it Age: Isochrones}---The color and absolute magnitude of
TIC~141146667 suggest that it is a pre-main-sequence M dwarf, similar
to all other known CPVs \cite{Stauffer2017,Stauffer2021,Bouma2024}.
The star's proximity ($d$=58\,pc) and its high galactic latitude
($b$=$+$53$^\circ$) yield negligible interstellar reddening along the
line of sight \cite{Green2019}.
{\bf Figure~X} shows the location of the star in the color--absolute
magnitude diagram (CAMD)
relative to young stellar populations including Upper Scorpius (USco),
Upper-Centaurus-Lupus (UCL), IC~2602, and the Pleiades.
To make this diagram, {\bf TODO}we adopted the USco and UCL members
from X, the IC~2602 members from Y, and the Pleiades members from Z,
and applied the same reddening corrections as in \cite{Bouma2022}.
To measure the isochrone age of TIC~141146667, we then followed the 
empirical approach proposed by \cite{Gagne2020} and
implemented in \cite{Bouma2022}.
The idea is to estimate the star (or star cluster's) age by
interpolating between the observed isochrones of star clusters with
ages that have been otherwise calibrated, for instance using
the lithium depletion-boundary method.
While this circumvents known modelling issues in molecular opacities
and starspot coverage for young M dwarfs,
the ages returned by this procedure depend on the ages
assumed for each reference cluster.
In this case, the most important reference is UCL, since TIC~141146667
overlaps directly on its isochrone on the CAMD.
We adopted the 16~Myr age for UCL from \cite{PecautMamajek2016}, which
is consistent with ages determined by more recent sub-clustering
analyses \cite{Ratzenbock2023} for the largest sub-populations.
This in turns yields an isochronal age for TIC~141146667 of
$t_{\rm iso}$=$16^{+19}_{-6}$\,Myr.
Qualitatively, this range of uncertainty is consistent with no stars
in the $\approx$40\,Myr IC~2602 cluster with similar redness as
TIC~141146667 being equally bright.
%FIXME: update discussion following consideration of anomalous CPV
% brightnesses and colors...

{\it Age: Lithium}---The depletion of lithium due to $^7$Li(p,
$\alpha$)$^4$He burning in the cores of low-mass stars has been
studied for over sixty years
\cite{Hayashi1963,Bildsten1997,Burke2004}.  \cite{Wood2023} provided
an instructive recent overview: an abbreviated summary is that the M
dwarfs that have not yet had time to burn their photospheric lithium
show the lithium 6707.8\,\AA doublet in absorption, $\gtrsim$10\%
below their ``continua''.  Relative to Sun-like stars, defining the
continuum for cool M dwarfs requires care due to their complex
spectra.  To attempt a lithium measurement, we constructed a
wavelength-binned and Doppler-corrected TIC~141146667 spectrum,
assigned its uncertainties based on the measured scatter across the
five hour dataset, and compared against the nearest matching template
from \cite{Bochanski2007}.  The data show a slight depression near the
expected lithium wavelength, potentially consistent with a $\Delta
\lambda$$\approx$3\,\AA line broadening.  This feature nominally
yields ${\rm EW}_{Li}$=71$^{+18}_{13}$\,m\AA, where the statistical
uncertainties are evaluated using a bootstrap resampling technique
from the statistical uncertainties in the spectrum.  Unfortunately,
the systematic uncertainties associated with the continuum
normalization are large, comparable to the amplitude of the feature;
we therefore prefer to quote the results of this measurement as a
$2\sigma$ upper limit: ${\rm EW}_{Li}$$<$107\,m\AA.

While

W

Frankly, it's not obviously there!  (but maybe
this is bc of the blending, and a pEW would show smth small)
%TODO TODO TODO
